Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{ Funded by DARPA Cyber Assured Systems Engineering (CASE) contract HR00111890001. }

### 2021-05-11 paper/agree.tex
Update with Hardin feedback

\begin{figure*}
  \begin{center}
    \begin{tabular}{c}
      \includegraphics[scale=0.4]{example.png} \\
      (a) \\
      \includegraphics[scale=0.4]{example-certificate.png} \\
      (b)
    \end{tabular}
  \end{center}
\caption{Automated UAV route planning system. (a) Unhardened system. (b) Failure certificate.}
\label{fig:example}
\end{figure*}

\figref{fig:example} is an AADL description of an implementation of a software system for route planning and automated control for a UAV. It is loosely based on the system in the case study introduced later. The system receives an automation request that is forwarded to an untrusted third-party route planner (AI) that decides the flight path of the UAV based on its current position and the requested task. The waypoint manager (WM) receives the mission command as a set of waypoints from the planner and starts the UAV flying the mission, issuing waypoints to the UAV flight controller as the UAV location changes. It is an \emph{as is} legacy component. The alert manager never raises the alert as there is no capacity in the design to know when an alert is needed.

The goal is to cyber-harden the system so that it is resilient to cyber-attack. For this simple example, two attacks are considered; both attacks originating from the untrusted route-planner: first, malformed messages to disrupt the waypoint manager or UAV flight controller; and second, a delayed or unrequested automation response to prevent the UAV from flying a mission or to fly the wrong mission.

@@

eq today : bool = (req = rsp);
eq yesterday : bool = (req and not rsp) ->
                      pre(req and not rsp) and (not req and rsp);
eq policy : bool = today or yesterday;
eq since : bool =  alrt or (alrt and (false -> pre(since)));

@@

This work uses an assume guarantee paradigm, in the form of contracts, to model and hierarchically prove properties of the system (cite AGREE). The contracts constrain the inputs and state properties of output for component models. In this example, the model for the route planner assumes its input is well-formed, but it makes no guarantees about its output as it is an untrusted third party component. The contract for the legacy waypoint manager makes similar assumptions about its input, but unlike the route planner, the model constrains its output to be well-formed and coincide with the mission command from the route planner and the air vehicle location.\footnote{See \cite{repo} for the full model.}

The contract to model the SW component with its primary inputs and outputs is in \figref{fig:sw}. The contract language is a first order predicate calculus. The contract semantics are synchronous data-flow where the inputs, outputs, and expressions are characterized by data streams. The semantics are such that the contracts are evaluated in dependency order with inputs being propagated to outputs through all the contracts until they stabilize; as such, the contracts, and thereby the system model, must be acyclic. Once the contracts have stabilized, then the model takes a synchronous step to the next input data in the stream. The semantics do not model computation or communication delay. The output of one contract is seen at the input of any downstream contract in the same step of the input data stream.

@@

There are two requirements in the SW model, among others, related to the cyber-attacks: the first, \emph{Waypoint is well-formed} disallows malformed waypoints from being propagated; and the second, \emph{Alert if start is not bounded relative to a request}, disallows a mission being started without a request or if the start is delayed more than one step after a request. This later guarantee merits some further discussion as it illustrates the steam nature of the semantics.

The guarantee is an invariant on the expression \texttt{policy or since} meaning that either the policy holds or the alert is sounding. The \texttt{policy} is defined by two local values: \texttt{today} and \texttt{yesterday}. The \texttt{today} value is asserted when in the current time step there is a request with a response, or there is no request and no response.  The value of \texttt{yesterday} in the current time step relies on values from the previous time step \texttt{->} operator for initialization as the previous time step is undefined in the first step of the system. The left operand is the initial value of \texttt{yesterday} at start, which in this example is \texttt{(req and not rsp)}, because seeing a request with no response is inconclusive in the first step of the system. The right operand is the value of \texttt{yesterday} after the initial step. Here the \texttt{pre} operator refers to the value of the expression \texttt{(req and not rsp)} in the prior time step. Intuitively, \texttt{yesterday} is true if the previous time step made a request without a matching response and the current time step has the matching response to that request with no new request.

@@

Model checking proves whether the contract models of the components, with their connections, implement the system model in \figref{fig:sw}. The results of the model checker are in \figref{fig:example}(b). The red exclamation points are properties that do not hold. Each comes with a counter-example. The result is not unexpected since the model of the route planner leaves the output behavior unconstrained (i.e., an untrusted component can do anything). The counter-example for \emph{Alert if start is not bounded relative to a request} shows a response in the first time step with no matching request.

\begin{figure*}
  \begin{center}
    \begin{tabular}{c}
      \includegraphics[scale=0.3]{hardened.png} \\
      (a) \\
      \includegraphics[scale=0.4]{hardened-certificate.png} \\
      (b)
    \end{tabular}
  \end{center}
  \caption{Hardened UAV system. (a) The implementation with high-assurance components. (b) Passing certificate.}
  \label{fig:hardened}
\end{figure*}

@@

\newsavebox{\mntr}
\begin{lrbox}{\mntr}
\begin{lstlisting}[style=agree]
const is_latched : bool = 
  Get_Property(this, CASE_Properties::Monitor_Latched);
eq rsp : bool = event(Response);
eq req : bool = event(Request);
eq today : bool = (req = rsp);
eq yesterday : bool = (req and not rsp) ->
                      pre(req and not rsp) and (not req and rsp);
eq policy : bool = today or yesterday;
eq alert : bool = (not policy) 
               -> ((is_latched and pre(alert)) or not policy);
guarantee Monitor_Alert 
  "Alert port tracks alert variable" :
  event(Alert) = alert;
guarantee Monitor_Output
  "Output if not alerted" :
  if event(Alert) then (not event(Output)) else
  if event(Response) then (event(Output) and (Output = Response))
  else (not event(Output));
\end{lstlisting}
\end{lrbox}

@@

The system implementation is cyber-hardened by using the CASE tools to insert high-assurance components in the form of a filter and a monitor as shown in \figref{fig:hardened}(a). A filter enforces an invariant over each datum in the data stream by not forwarding input to its output if that input is violating. Its stylized form as generated by the tool is shown in \figref{fig:assurance}(a). The user must declare the meaning of the \texttt{policy} value as part of the transform on the model. In this example, \texttt{policy} is the assumption made by the waypoint manager about the automation response being well-formed. The output guarantees are automatically generated by the tool.

A monitor captures a relation over time on input data and is able to reason about temporal properties of that input. By definition, it raises an alert output when the relation is violated. The generated monitor contract is shown in \figref{fig:assurance}(b). The \texttt{is\_latched} value makes the alert persistent meaning that once raised always raised. The definition for \texttt{policy} is taken from the contract in \figref{fig:sw}. As before the guarantees for the outputs are autogenerated by the tool.

The proof certificate for the new cyber-hardened implementation is in \figref{fig:hardened}(b). The high-assurance components guarantee the correct behavior of the SW implementation in the presence of the considered cyber attacks.

\newsavebox{\cml}
\begin{lrbox}{\cml}
\begin{lstlisting}[style=myML]
fun filter_step () =
let val () = Utils.clear_buf buffer
    val () = API.callFFI "get_input" "" buffer
    val string = Utils.buf2string buffer
    val policy = WELL_FORMED_AUTOMATION_RESPONSE string 
in
    if policy
    then
        let val string = Word8Array.substring buffer 1
                         (Word8Array.length buffer - 1)
        in API.callFFI "put_output" string Utils.emptybuf
        end
    else print"Filter rejects message.\n"
end
\end{lstlisting}
\end{lrbox}

\begin{figure}
  \begin{center}
    \begin{tabular}{c}
      \scalebox{0.60}{\usebox{\cml}}
    \end{tabular}
  \end{center}
  \caption{Synthesized CakeML for the filter.}
  \label{fig:cakeml}
\end{figure}

The high-assurance components are automatically synthesized from the contract models to equivalent models in the CakeML language. The synthesized CakeML code for the filter is shown in \figref{fig:cakeml}. The code is called at dispatch by the scheduler. The \texttt{API.callFFI} is the link to the communication fabric to capture input and provide output. The body of the function restates the filter contract to make the appropriate assignments in a way that matches the truth value of the predicate in the filter guarantee.

CakeML itself provides a complete verified compilation to binaries for several different platforms meaning that the resulting binaries exactly preserve the meaning of the original CakeML code. A similar proof is given for the synthesis of the contract model for a high-assurance component to CakeML. A high-assurance component contract has a precise meaning in terms of data streams, and the synthesis exactly preserves that meaning in the resulting CakeML. In other words, for any set of input streams that meet the component's contract assumptions, the output streams from the synthesized CakeML exactly match the output streams from the high-assurance component's contract.

Preserving the input/output relationship of streams between the two models lifts the contract verification results to the deployed system. If the contract model verifies, then the meaning of those results hold for the deployed system if the other components implement their contracts, an appropriate schedule exists that follows the dependent data-flow, and the communication fabric works as expected.

>>>>

@@

eq current : bool = (req = rsp);
eq previous : bool = (req and not rsp) ->
                      pre(req and not rsp) and (not req and rsp);
eq policy : bool = current or previous;
eq since : bool =  alrt or (alrt and (false -> pre(since)));

@@

The revised AGREE specification for the SW component with its primary inputs and outputs is in \figref{fig:sw}. Without going into too much detail, the language is a first order predicate calculus. The semantics are synchronous data-flow where the inputs, outputs, and expressions are characterized by data streams. The semantics are such that the contracts are evaluated in dependency order with inputs being propagated to outputs through all the contracts until they stabilize; as such, the contracts, and thereby the system model, must be acyclic. Once the contracts have stabilized, then the model takes a synchronous step to the next input data in the stream. The semantics do not model computation or communication delay. The output of one contract is seen at the input of any downstream contract in the same step of the input data stream.

@@

The guarantee is an invariant on the expression \texttt{policy or since} meaning that either the policy holds or the alert is sounding. The \texttt{policy} is defined by two local values: \texttt{current} and \texttt{previous}. The \texttt{current} value is asserted when in the current time step there is a request with a response, or there is no request and no response.  The value of \texttt{previous} in the current time step relies on values from the previous time step \texttt{->} operator for initialization as the previous time step is undefined in the first step of the system. The left operand is the initial value of \texttt{previous} at start, which in this example is \texttt{(req and not rsp)}, because seeing a request with no response is inconclusive in the first step of the system. The right operand is the value of \texttt{previous} after the initial step. Here the \texttt{pre} operator refers to the value of the expression \texttt{(req and not rsp)} in the prior time step. Intuitively, \texttt{previous} is true if the previous time step made a request without a matching response and the current time step has the matching response to that request with no new request.

@@

@@

\newsavebox{\mntr}
\begin{lrbox}{\mntr}
\begin{lstlisting}[style=agree]
const is_latched : bool = 
  Get_Property(this, CASE_Properties::Monitor_Latched);
eq rsp : bool = event(Response);
eq req : bool = event(Request);
eq current : bool = (req = rsp);
eq previous : bool = (req and not rsp) ->
                      pre(req and not rsp) and (not req and rsp);
eq policy : bool = current or previous;
eq alert : bool = (not policy) 
                -> ((is_latched and pre(alert)) or not policy);
guarantee Monitor_Alert 
  "Alert port tracks alert variable" :
  event(Alert) = alert;
guarantee Monitor_Output
  "Output if not alerted" :
  if event(Alert) then (not event(Output)) else
  if event(Response) then (event(Output) and (Output = Response))
  else (not event(Output));
\end{lstlisting}
\end{lrbox}

@@

The hardened system is implemented by inserting the high-assurance components. The AGREE specifications for both are in \figref{fig:hardened}. Both are stylized for synthesis. The stylized filter form as generated by the tool is shown in \figref{fig:assurance}(a). The user must declare the meaning of the \texttt{policy} value as part of the transform on the model. In this example, \texttt{policy} is the assumption made by the waypoint manager about the automation response being well-formed. The output guarantees are automatically generated by the tool. The monitor is shown in \figref{fig:assurance}(b). The \texttt{is\_latched} value makes the alert persistent meaning that once raised always raised. The definition for \texttt{policy} is taken from the contract in \figref{fig:sw}. As before the guarantees for the outputs are autogenerated by the tool.

Add text to this section that defines the meaning of the contract over streams. State the prove guarantee from AGREE is terms of the streams. The statement will be used in the synthesis section to argue correctness.

### 2021-05-11 paper/example.tex
Update with Hardin feedback

\figref{fig:example} is an AADL description of an implementation of a software system for route planning and automated control for a UAV. It is loosely based on the system in the case study introduced later. The system receives an automation request that is forwarded to an untrusted third-party route planner (AI) that decides the flight path of the UAV based on its current position and the requested task. The waypoint manager (WM) receives the mission command as a set of waypoints from the planner and starts the UAV flying the mission, issuing waypoints to the UAV flight controller as the UAV location changes. It is an \emph{as is} legacy component. The alert manager never raises the alert as there is no capacity in the design to know when an alert is needed.

The goal is to cyber-harden the system so that it is resilient to cyber-attack. For this simple example, two attacks are considered; both attacks originating from the untrusted route-planner: first, malformed messages to disrupt the waypoint manager or UAV flight controller; and second, a delayed or unrequested automation response to prevent the UAV from flying a mission or to fly the wrong mission.

\newsavebox{\sw}
\begin{lrbox}{\sw}
\begin{lstlisting}[style=agree]
eq req : bool = event(AutomationRequest);
eq avl : bool = event(AirVehicleLocation);
eq wp : bool = event(Waypoint);
eq rsp: bool = event(Start);
eq alrt : bool = event(Alert);

assume "Automation request is well-formed" :
    req => WELL_FORMED_AUTOMATION_REQUEST(AutomationRequest);
assume "Air vehicle location is well-formed" :
    avl => WELL_FORMED_WAYPOINT(AirVehicleLocation);

eq today : bool = (req = rsp);
eq yesterday : bool = (req and not rsp) ->
                      pre(req and not rsp) and (not req and rsp);
eq policy : bool = today or yesterday;
eq since : bool =  alrt or (alrt and (false -> pre(since)));

guarantee "Start includes a waypoint" :
    rsp => wp;
guarantee "Locations required after the start waypoint" :
    (wp and not rsp) => avl;
guarantee "Waypoint is well-formed" :
    wp => WELL_FORMED_WAYPOINT(Waypoint);
guarantee "Alert if start is not bounded relative to a request" :
    policy or since;
\end{lstlisting}
\end{lrbox}

\begin{figure}
  \begin{center}
    \scalebox{0.60}{\usebox{\sw}}
  \end{center}
  \caption{The SW component contract.}
  \label{fig:sw}
\end{figure}

This work uses an assume guarantee paradigm, in the form of contracts, to model and hierarchically prove properties of the system (cite AGREE). The contracts constrain the inputs and state properties of output for component models. In this example, the model for the route planner assumes its input is well-formed, but it makes no guarantees about its output as it is an untrusted third party component. The contract for the legacy waypoint manager makes similar assumptions about its input, but unlike the route planner, the model constrains its output to be well-formed and coincide with the mission command from the route planner and the air vehicle location.\footnote{See \cite{repo} for the full model.}

The contract to model the SW component with its primary inputs and outputs is in \figref{fig:sw}. The contract language is a first order predicate calculus. The contract semantics are synchronous data-flow where the inputs, outputs, and expressions are characterized by data streams. The semantics are such that the contracts are evaluated in dependency order with inputs being propagated to outputs through all the contracts until they stabilize; as such, the contracts, and thereby the system model, must be acyclic. Once the contracts have stabilized, then the model takes a synchronous step to the next input data in the stream. The semantics do not model computation or communication delay. The output of one contract is seen at the input of any downstream contract in the same step of the input data stream.

The contract in \figref{fig:sw} uses \texttt{eq} statements to define variables local to the contract. For example, the \texttt{req} variable is equivalent to the \texttt{event(AutomationRequest)} expression. There is an implicit \emph{event} input (or output) associated with every named event port in a component. The semantics used here do not not buffer these events so the implicit input (or output) is only a boolean value. An \texttt{event} expression refers to that implicit input (or output) and is true if there is data on the named port. The system contract assumes well-formed input and guarantees a set of properties about the output.

There are two requirements in the SW model, among others, related to the cyber-attacks: the first, \emph{Waypoint is well-formed} disallows malformed waypoints from being propagated; and the second, \emph{Alert if start is not bounded relative to a request}, disallows a mission being started without a request or if the start is delayed more than one step after a request. This later guarantee merits some further discussion as it illustrates the steam nature of the semantics.

The guarantee is an invariant on the expression \texttt{policy or since} meaning that either the policy holds or the alert is sounding. The \texttt{policy} is defined by two local values: \texttt{today} and \texttt{yesterday}. The \texttt{today} value is asserted when in the current time step there is a request with a response, or there is no request and no response.  The value of \texttt{yesterday} in the current time step relies on values from the previous time step \texttt{->} operator for initialization as the previous time step is undefined in the first step of the system. The left operand is the initial value of \texttt{yesterday} at start, which in this example is \texttt{(req and not rsp)}, because seeing a request with no response is inconclusive in the first step of the system. The right operand is the value of \texttt{yesterday} after the initial step. Here the \texttt{pre} operator refers to the value of the expression \texttt{(req and not rsp)} in the prior time step. Intuitively, \texttt{yesterday} is true if the previous time step made a request without a matching response and the current time step has the matching response to that request with no new request.

The value of \texttt{since} in \emph{Alert if start is not bounded relative to a request} relies on its own value in the previous time step. The intuitive reading of the expression is that the alert has been true since the time when it first sounded. The first \texttt{alrt} sets \texttt{since} to true, and once the value of \texttt{since} is true, that value persists as long as \texttt{alrt} holds. The \emph{Alert if start is not bounded relative to a request} guarantee defines one behavior of a correctly implemented system. Together with the other guarantees, the contract models the expected input and output of the system as a whole.

Model checking proves whether the contract models of the components, with their connections, implement the system model in \figref{fig:sw}. The results of the model checker are in \figref{fig:example}(b). The red exclamation points are properties that do not hold. Each comes with a counter-example. The result is not unexpected since the model of the route planner leaves the output behavior unconstrained (i.e., an untrusted component can do anything). The counter-example for \emph{Alert if start is not bounded relative to a request} shows a response in the first time step with no matching request.

@@

\newsavebox{\flt}
\begin{lrbox}{\flt}
\begin{lstlisting}[style=agree]
eq policy : bool = 
  WELL_FORMED_AUTOMATION_RESPONSE(Input);       
guarantee Filter_Output "Filter output is well-formed" :
  if event(Input) and policy then 
    event(Output) and Output = Input
  else not event(Output);
\end{lstlisting}
\end{lrbox}

\newsavebox{\mntr}
\begin{lrbox}{\mntr}
\begin{lstlisting}[style=agree]
const is_latched : bool = 
  Get_Property(this, CASE_Properties::Monitor_Latched);
eq rsp : bool = event(Response);
eq req : bool = event(Request);
eq today : bool = (req = rsp);
eq yesterday : bool = (req and not rsp) ->
                      pre(req and not rsp) and (not req and rsp);
eq policy : bool = today or yesterday;
eq alert : bool = (not policy) 
               -> ((is_latched and pre(alert)) or not policy);
guarantee Monitor_Alert 
  "Alert port tracks alert variable" :
  event(Alert) = alert;
guarantee Monitor_Output
  "Output if not alerted" :
  if event(Alert) then (not event(Output)) else
  if event(Response) then (event(Output) and (Output = Response))
  else (not event(Output));
\end{lstlisting}
\end{lrbox}

\begin{figure}
  \begin{center}
    \begin{tabular}{c}
      \scalebox{0.60}{\usebox{\flt}} \\
      (a) \\
      \scalebox{0.60}{\usebox{\mntr}} \\
      (b)
    \end{tabular}
  \end{center}
  \caption{High-assurance component contracts. (a) The filter. (b) The monitor.}
  \label{fig:assurance}
\end{figure}

The system implementation is cyber-hardened by using the CASE tools to insert high-assurance components in the form of a filter and a monitor as shown in \figref{fig:hardened}(a). A filter enforces an invariant over each datum in the data stream by not forwarding input to its output if that input is violating. Its stylized form as generated by the tool is shown in \figref{fig:assurance}(a). The user must declare the meaning of the \texttt{policy} value as part of the transform on the model. In this example, \texttt{policy} is the assumption made by the waypoint manager about the automation response being well-formed. The output guarantees are automatically generated by the tool.

A monitor captures a relation over time on input data and is able to reason about temporal properties of that input. By definition, it raises an alert output when the relation is violated. The generated monitor contract is shown in \figref{fig:assurance}(b). The \texttt{is\_latched} value makes the alert persistent meaning that once raised always raised. The definition for \texttt{policy} is taken from the contract in \figref{fig:sw}. As before the guarantees for the outputs are autogenerated by the tool.

The proof certificate for the new cyber-hardened implementation is in \figref{fig:hardened}(b). The high-assurance components guarantee the correct behavior of the SW implementation in the presence of the considered cyber attacks.

\newsavebox{\cml}
\begin{lrbox}{\cml}
\begin{lstlisting}[style=myML]
fun filter_step () =
let val () = Utils.clear_buf buffer
    val () = API.callFFI "get_input" "" buffer
    val string = Utils.buf2string buffer
    val policy = WELL_FORMED_AUTOMATION_RESPONSE string 
in
    if policy
    then
        let val string = Word8Array.substring buffer 1
                         (Word8Array.length buffer - 1)
        in API.callFFI "put_output" string Utils.emptybuf
        end
    else print"Filter rejects message.\n"
end
\end{lstlisting}
\end{lrbox}

\begin{figure}
  \begin{center}
    \begin{tabular}{c}
      \scalebox{0.60}{\usebox{\cml}}
    \end{tabular}
  \end{center}
  \caption{Synthesized CakeML for the filter.}
  \label{fig:cakeml}
\end{figure}

The high-assurance components are automatically synthesized from the contract models to equivalent models in the CakeML language. The synthesized CakeML code for the filter is shown in \figref{fig:cakeml}. The code is called at dispatch by the scheduler. The \texttt{API.callFFI} is the link to the communication fabric to capture input and provide output. The body of the function restates the filter contract to make the appropriate assignments in a way that matches the truth value of the predicate in the filter guarantee.

CakeML itself provides a complete verified compilation to binaries for several different platforms meaning that the resulting binaries exactly preserve the meaning of the original CakeML code. A similar proof is given for the synthesis of the contract model for a high-assurance component to CakeML. A high-assurance component contract has a precise meaning in terms of data streams, and the synthesis exactly preserves that meaning in the resulting CakeML. In other words, for any set of input streams that meet the component's contract assumptions, the output streams from the synthesized CakeML exactly match the output streams from the high-assurance component's contract.

>>>>

\figref{fig:example} is an diagram form an AADL model of an implementation of a software system  (SW) for route planning and automated control for a UAV. It is loosely based on the system in the case study introduced later. The system receives an automation request that is forwarded to an untrusted third-party route planner (AI) that decides the flight path of the UAV based on its current position and the requested task. The waypoint manager (WM) receives the mission command as a set of waypoints from the planner and starts the UAV flying the mission, issuing waypoints to the UAV flight controller as the UAV location changes. It is an \emph{as is} legacy component.

The expected behavior of the SW system, and the components in its implementation, are modeled with AGREE contracts. The contracts constrain input and state properties of output for component models. AGREE model checks this assume-guarantee system to hierarchically prove the composite system obeys all contract obligations under all possible finite input streams.

The initial AGREE specification for the components and SW system make no assumptions about the integrity of the inputs and outputs. The cyber-vulnerability analysis identifies the potential of the untrusted AI route planner component to behave maliciously. As such, the AGREE specification for the AI route planner is updated to model this ability to behave maliciously as an untrusted component by removing any guarantees about its output; in other words, its output is unconstrained in the AGREE specification allowing it to take on any value and allowing it to send that value at any time.

The contracts for the other components are also updated with the threat analysis information. For example, the specification for the legacy waypoint manager adds assumptions about about its input being \emph{well-formed} since that is no longer known \emph{a priori} as AI route planner is now untrusted.

There are two requirements added to the SW AGREE specification related to the cyber-vulnerability analysis: the first, \emph{Waypoint is well-formed} disallows malformed waypoints from being propagated downstream; and the second, \emph{Alert if start is not bounded relative to a request}, disallows a mission being started without a request or if the start is delayed more than one step after a request. Here the goal is to make clear that a delayed or unrequested automation response may prevent the UAV from flying its intended mission or to fly the wrong mission.

AGREE model checks the composition of the new specifications for the implementation in \figref{fig:example}(a) to see if it complies with the added cyber-resilience properties. The results of the model checker are in \figref{fig:example}(b). The red exclamation points are properties that do not hold. Each comes with a counter-example. The results are not unexpected given the AI route planner's unconstrained behavior and the new assumptions about input to the legacy waypoint manager. The counter-example for \emph{Alert if start is not bounded relative to a request} shows a response in the first time step with no matching request. A clear malicious behavior from the untrusted component.

@@

The system implementation is cyber-hardened by automatically transforming the model to insert high-assurance components in the form of a filter and a monitor as shown in \figref{fig:hardened}(a). A filter enforces an invariant over each datum in the data stream by not forwarding input to its output if that input is violating. The auto-generated AGREE specification states that only well-formed inputs are passed to the output. The user must provide this filtering policy, but it is usually based on the existing assumptions made by downstream components that consume the filter output.

A monitor captures a relation over time on input data and is able to reason about temporal properties of that input. It raises an alert if ever the specified temporal properties are violated. The auto-generated AGREE specification for the monitor states that an automation response can only be generated in conjunction with an automation request, and that response must come with the request or in the next step after the request. As with the filter, the user provides the policy, and that policy is based on the existing AGREE specification in the SW system.

The AGREE analysis of the cyber-hardened implementation with the auto-generated high-assurance components is shown in \figref{fig:hardened}(b). Here AGREE gives a proof certificate that the high-assurance components guarantee the correct behavior of the SW implementation in the presence of the considered cyber vulnerability from the untrusted AI route planner.

The high-assurance components are automatically synthesized from the AGREE specifications to equivalent models in the CakeML language. CakeML itself provides a complete verified compilation to binaries for several different platforms meaning that the resulting binaries exactly preserve the meaning of the original CakeML code.

A similar proof is given for the synthesis of the contract model for a high-assurance component to CakeML. A high-assurance component contract has a precise meaning in terms of data streams, and the synthesis exactly preserves that meaning in the resulting CakeML. In other words, for any set of input streams that meet the component's contract assumptions, the output streams from the synthesized CakeML exactly match the output streams from the high-assurance component's contract.

### 2021-05-11 paper/intro.tex
Update with Hardin feedback

Here is the basic process:
\begin{enumerate}
  \item CakeML provides a mechanism to take a defined logic in HOL4 and turn it into CakeML in a way that exactly preserves the meaning of the logic.
  \item Verified components are created by defining the meaning of the components in HOL4, and then using the existing theories to turn that meaning into CakeML
  \item Contiguity types provide a formalism for defining a language and recognizing membership in a language via a matcher
  \item Contiguity types include dependent types and the ability to define when a type is malformed
  \item Define a contract language that makes assumptions on input and guarantees properties on output
  \item Give meaning to the contract language in terms of traces (not easy to do because of the state)
  \item Show how the contract can be expressed in terms of a step function 
  \item The step-function is lifted to traces
  \item A contract defines a language over traces
  \item A step function recognizes membership in that language (prefix closed)
  \item Follow the same pattern used in contiguity types
  \item The CakeML comes for free once the HOL4 theories are shown
\end{enumerate}

@@

Motivate with some example: watchdog example on DNS monitoring for DOS attacks? Estimates the growth rate of the input request queue over time and sends a throttle signal if the growth is too fast (looks suspicious)---any bounded response of behavior based watch-dog is fine.

>>>>

In recent years, aerospace stakeholders have become aware that avionics systems are subject to possible cyber-attacks just like other cyber-physical systems.  In addition to being fault-tolerant, safety-critical avionics systems must also be {\em cyber-resilient}. Cyber-resiliency means that the system is tolerant to cyberattacks just as safety-critical systems are tolerant to random faults: they recover and continue to execute their mission function, or safely shut down, as requirements dictate.

Unfortunately, systems engineers are currently given few development tools to help answer even basic questions about potential vulnerabilities and mitigations, and instead rely on process-oriented checklists and guidelines.  Cyber vulnerabilities are often discovered during penetration testing late in the development process.  Worse yet, they may be discovered after the product has been fielded, necessitating extremely expensive and time-consuming remediation. This is not a sustainable development model.

The DARPA Cyber Assured Systems Engineering (CASE) program is targeted to developing design, analysis, and verification tools that enable systems engineers to {\em design-in} cyber-resiliency for complex cyber-physical systems. 
The result is a Model-Based Systems Engineering (MBSE) environment called {\em BriefCASE} which is based on the Architecture Analysis and Design Language (AADL)~\cite{aadl}.  BriefCASE extends the Open Source AADL Tool Environment (OSATE) to add new design, analysis, and code generation capabilities targeted at building cyber-resilient systems.

BriefCASE provides access to two analysis tools (GearCASE~\cite{gearcase2020} and DCRYPPS~\cite{dcrypps2019}) that can examine AADL models to detect potential cyber vulnerabilities and suggest requirements for mitigation.  
A library of architectural transforms guides systems engineers through automated model transformations that modify the architecture to address these requirements, possibly inserting new high-assurance components into the system. 
Implementations for these new high-assurance components are synthesized from formal specifications using the Semantic Properties for Language and Automata Theory (SPLAT) tool~\cite{slind-hcss2020}.
Formal verification that the transformed system model satisfies its cyber requirements is accomplished via model checking using the Assume Guarantee Reasoning Environment (AGREE)~\cite{agree2013}.  Cyber-resilient code implementing the verified model is automatically generated using the High Assurance Modeling and Rapid Engineering for Embedded Systems (HAMR) toolkit~\cite{hamr}.  If desired, this code can be targeted to the formally verified seL4 secure microkernel~\cite{sel4-2009}.

A novel aspect of the approach is the ability to automatically transformation the model to insert high-assurance components to harden the system against potential cyber-vulnerabilities. The behavior of the model is given by contracts on components expressed in the AGREE modeling language. These contracts are modified as a result of the cyber-vulnerability analysis to assume properties on the input and and guarantee properties of the output that reflect a hardened system. In the absence of any transformation to implement the hardening, AGREE fails to prove the system cyber-hardened.

The two automatic transforms discussed in this paper are the insertion of filters to prevent malformed data from malicious actors from being propagated to downstream components, and monitors to detect and alert unexpected behaviors arising from untrusted components. These transformations not only change the architecture of the model by adding in new components, but they generate a complete formal specification for the behavior of the inserted high-assurance components in the AGREE language. Those specifications are sufficient for model checking to prove that with the high-assurance components, hardened system meets its cyber-resiliency requirements.

The other novel aspect of the approach is the synthesis of the AGREE specifications for the high-assurance components to CakeML. This paper agues that synthesis to CakeML to be correct. CakeML then provides a verified compilation path to several different target binaries proving that the meaning of the CakeML is exactly preserved in the final binaries. Assuming the deployed cyber-hardened system is scheduled as intended by the AADL model, and the HAMR generated communication fabric delivers messages between components as expected, then the AGREE model checking results hold for the deployed system in that in detects and prevents the indicated cyber-vulnerabilities over all possible finite inputs. Future work is lifting this result to infinite input traces as these systems are inherently reactive and intended to run forever.

The approach is motivated, and illustrated, in a simple example. A full case study applying these transformations to a UAV system that uses the Air Force Research Laboratory's OpenUxAS services for route planning is briefly reported. Here the transforms add filters to guard against malformed input and monitors to guard against ground station spoofing and malicious flight plans from OpenUxAS. The case study system is significantly more complex than the example and shows the viability of the modeling approach to full-scale industrial design.

\begin{comment}
Here is the basic process:
\begin{enumerate}
  \item CakeML provides a mechanism to take a defined logic in HOL4 and turn it into CakeML in a way that exactly preserves the meaning of the logic.
  \item Verified components are created by defining the meaning of the components in HOL4, and then using the existing theories to turn that meaning into CakeML
  \item Contiguity types provide a formalism for defining a language and recognizing membership in a language via a matcher
  \item Contiguity types include dependent types and the ability to define when a type is malformed
  \item Define a contract language that makes assumptions on input and guarantees properties on output
  \item Give meaning to the contract language in terms of traces (not easy to do because of the state)
  \item Show how the contract can be expressed in terms of a step function 
  \item The step-function is lifted to traces
  \item A contract defines a language over traces
  \item A step function recognizes membership in that language (prefix closed)
  \item Follow the same pattern used in contiguity types
  \item The CakeML comes for free once the HOL4 theories are shown
\end{enumerate}

@@

Motivate with some example: watchdog example on DNS monitoring for DOS attacks? Estimates the growth rate of the input request queue over time and sends a throttle signal if the growth is too fast (looks suspicious)---any bounded response of behavior based watch-dog is fine.
\end{comment}

### 2021-05-11 paper/synthesis.tex
Update with Hardin feedback

>>>>

\newsavebox{\cml}
\begin{lrbox}{\cml}
\begin{lstlisting}[style=myML]
fun filter_step () =
let val () = Utils.clear_buf buffer
    val () = API.callFFI "get_input" "" buffer
    val string = Utils.buf2string buffer
    val policy = WELL_FORMED_AUTOMATION_RESPONSE string 
in
    if policy
    then
        let val string = Word8Array.substring buffer 1
                          (Word8Array.length buffer - 1)
        in API.callFFI "put_output" string Utils.emptybuf
        end
    else print"Filter rejects message.\n"
end
\end{lstlisting}
\end{lrbox}

\begin{figure}
  \begin{center}
    \begin{tabular}{c}
      \scalebox{0.60}{\usebox{\cml}}
    \end{tabular}
  \end{center}
  \caption{Synthesized CakeML for the filter.}
  \label{fig:cakeml}
\end{figure}

The synthesized CakeML code for the filter is shown in \figref{fig:cakeml}. The code is called at dispatch by the scheduler. The \texttt{API.callFFI} is the link to the communication fabric to capture input and provide output. The body of the function restates the filter contract to make the appropriate assignments in a way that matches the truth value of the predicate in the filter guarantee.
The auto-generated AGREE specification raises an alert output when the relation is violated.
